---
title: "Listing the functions of the netCDF libraries"
author: "Fadi Moubayed"
date: "9/24/2020"
output: html_document
---

### Loading packages
```{r message=FALSE,warning=FALSE}
# Available netCDF libraires


require(ncdf4)
require(ncdf4.helpers)
require(tidync)
require(RNetCDF)
require(ncdfgeom)
require(stars)
```


### The following function lists the functions available in the loaded libraries
Extra libraries are also being listed using this function
```{r}
ListLoadedLibrariesFunctions = function(){
  # Listing the default libraries. These do not need to be listed
  defaultLibraries = list("stats","graphics" ,"grDevices","utils","datasets" ,"methods","base")
  # Listing the loaded libraries
  loadedLibraries = (.packages())
  # Getting the loaded libraries only
  LoadedlibrariesOnly = setdiff(loadedLibraries, defaultLibraries)
  
  # Adding package: to the names of the libraries
  libraries1= paste0("package:", LoadedlibrariesOnly)
  sapply(libraries1, ls)
}

```

```{r}
#ListLoadedLibrariesFunctions()
```

### Adding the libraries that need to be listed maually
The libraries are added manually to a list and then the functions are listed.
The libararies must be loaded otherwise they will not be listed and an error will be returned.
```{r}
libraries = c("ncdf4", "ncdf4.helpers", "tidync","RNetCDF", "ncdfgeom", "stars")
#adding package: to the names of the libraries
libraries1= paste0("package:", libraries)
sapply(libraries1, ls)
```

