---
title: "Listing functions of GeoJSON libraries in R"
author: "Fadi Moubayed"
date: "2/01/2021"
output: html_document
---

### Loading packages
```{r message=FALSE,warning=FALSE}
library(geojsonio)
library(geojson)
library(geojsonlint)
library(geojsonR)
library(wellknown) # for adding properties to GeoJson files
```


### The following function lists the functions available in the loaded libraries
Extra libraries are also being listed using this function
```{r}
ListLoadedLibrariesFunctions = function(){
  # Listing the default libraries. These do not need to be listed
  defaultLibraries = list("stats","graphics" ,"grDevices","utils","datasets" ,"methods","base")
  # Listing the loaded libraries
  loadedLibraries = (.packages())
  # Getting the loaded libraries only
  LoadedlibrariesOnly = setdiff(loadedLibraries, defaultLibraries)
  
  # Adding package: to the names of the libraries
  libraries1= paste0("package:", LoadedlibrariesOnly)
  sapply(libraries1, ls)
}

```

```{r}
ListLoadedLibrariesFunctions()
```

### Other functions that can be used from external libraries
#### st {sf}
Create simple feature from a numeric vector, matrix or list
Description
Create simple feature from a numeric vector, matrix or list

Usage
st_point(x = c(NA_real_, NA_real_), dim = "XYZ")

st_multipoint(x = matrix(numeric(0), 0, 2), dim = "XYZ")

st_linestring(x = matrix(numeric(0), 0, 2), dim = "XYZ")

st_polygon(x = list(), dim = if (length(x)) "XYZ" else "XY")

st_multilinestring(x = list(), dim = if (length(x)) "XYZ" else "XY")

st_multipolygon(x = list(), dim = if (length(x)) "XYZ" else "XY")

st_geometrycollection(x = list(), dims = "XY")
